A sheaf--function dictionary for algebraic tori over non-archimedean local fields

### 2013-08-16 shanghai.tex
Working on style in section 1

Let $\Frob{G}$ be the Frobenius automorphism of $\bG$.
Recall that $\Frob{G} = \id_{G}\times \Frob{\Spec{\bFq}}$ where
$\Frob{\Spec{\bFq}} = \Spec{\Frob{\Fq}}$ and $\Frob{\Fq}$  is the geometric Frobenius element in $\Gal(\bFq/\Fq)$.

@@

Morphisms in the category of quasicharacter sheaves on $G$, denoted by $\QC(G)$, are defined in the natural way:
\begin{enumerate}
\labitem{(CS.3)}{CS.3} if $\cs{L} = (\gcs{L},\mu,\phi)$ and
  $\cs{L'} = (\gcs{L'},\mu',\phi')$ are quasicharacter sheaves on $G$ then
  $\Hom(\cs{L},\cs{L'})$ is the set of morphisms $\alpha : \gcs{L} \to \gcs{L'}$
  of constructible $\ell$-adic sheaves on $\bG$ such that the following diagrams both commute.
  \[
  \begin{tikzcd}[column sep=40]
  \Frob{G}^* \gcs{L} \rar{\Frob{G}^* \alpha} \arrow[swap]{d}{\phi} & \Frob{G}^* \gcs{L'} \dar{\phi'}
  & %\arrow[draw=none]{d}[pos=.4,description]{\text{\normalsize{and}}} 
  & \arrow[swap]{d}{\mu} m^* \gcs{L} \rar{m^* \alpha} & m^* \gcs{L'} \dar{\mu'} \\
  \gcs{L} \rar{\alpha} & \gcs{L'}
  & {} & \LxL \rar{\tight{1}{\alpha\boxtimes \alpha}{1}} & \gcs{L'} \tight{0}{\boxtimes}{0} \gcs{L'}
  \end{tikzcd}
  \]
\end{enumerate}

@@

As mentioned after \cite{deligne:80a}*{Def.~1.1.10}, a Weil sheaf $(\gcs{L},\phi)$ on $G$ may be interpreted as a constructible $\ell$-adic sheaf $\gcs{L}$ on $\bG$ together with an action of the Weil group $\Weil{\Fq}$ on $\gcs{L}$ compatible with the action of $\Gal(\bFq/\Fq)$ on $\bG$. Since we will need this point of view later, we review that interpretation here. Setting 
\[
\varphi(\Frob{\Fq}^n) \ceq  \phi \circ \Frob{G}^*(\phi) \circ \cdots \circ (\Frob{G}^{n-1})^*(\phi)
\]
defines, for every $w\in \Weil{\Fq}$, an isomorphism
$\varphi(w) : w_G^* \gcs{L}\to \gcs{L}$, such that
$\varphi(uv) = \varphi(v) \circ v_G^* \varphi(u)$ for all $u,v\in \Weil{\Fq}$.
Note that $\phi = \varphi(\Frob{\Fq})$. 
If we further define $\varphi_1(w) \ceq (w_G)_*(\varphi(w)^{-1})$ then
$\varphi_1(w) : (w_G)_* \gcs{L}\to \gcs{L}$ is an isomorphism and
$\varphi_1(uv) = \varphi_1(u) \circ (u_G)_* \varphi_1(v)$ for all $u,v\in \Weil{\Fq}$.
This means the pair $(\gcs{L},\varphi_1)$ is almost an action $\Weil{\Fq}$ on $\gcs{L}$ compatible with the action of $\Gal(\bFq/\Fq)$ in the sense of \cite{deligne:SGA4.5}*{Expos\'e XIII,~1.1}. The pair $(\gcs{L},\varphi_1)$ fails to fulfill the requirements of \cite{deligne:SGA4.5}*{Expos\'e XIII,~1.1} only because $\Weil{\Fq}$ is profinite. This fact goes to the heart of the notion of a Weil sheaf and is one of the essential aspects of our definition of $\QC(G)$. 
%\end{remark}

@@

%\begin{proposition}\label{tensor}
The category $\QC(G)$ of quasicharacter sheaves on $G$ is rigid symmetric monoidal category \cite{etingof:09a} under the tensor product
$\cs{L} \otimes_{\EE}\cs{L'} \ceq (\gcs{L}\otimes\gcs{L'}, \mu\otimes\mu', \phi\otimes_{\EE}\phi')$
with duals given by applying the functor 
$\sheafHom(\ - \ ,\EE)$ (sheaf hom) to
conditions \ref{CS.0}, \ref{CS.1} and \ref{CS.2}.
%\end{proposition}
(The category of characters sheaves on $G$ is not abelian, so $\QC(G)$
is not a tensor category in the sense of \cite{deligne:02a}.)
%\begin{proof}\end{proof}
%
This rigid monoidal category structure for $\QC(G)$ puts a group
structure on the set $\QCiso{G}$ of isomorphism classes in
$\QC(G)$. Our main result regarding quasicharacter sheaves on $G$ (Theorem~\ref{thm:snake})
determines $\QCiso{G}$ under the hypothesis that the component group scheme for $G$ is geometrically finitely generated; see Theorem~\ref{thm:snake}. To do this, we use the trace of Frobenius.

@@

\section{Greenberg of Neron}
\label{sec:GN}

@@

This paper has shown how quasicharacter sheaves 
on the Greenberg transform of the Neron model of $T$ 
determine admissible characters of $T(K)$, and vice versa, functorially. When coupled with the reciprocity map $\text{rec}_T $ for $T$, 
which enjoys the same functorial properties, 
it follows that quasicharacter sheaves for $T$ 
determine Langlands parameters for $T$, and vice versa, functorially.
\[
\begin{tikzcd}
\ & \arrow[swap]{dl}{\trFrob{G}} \QCiso{\GN{T}} \arrow[dashed]{dr}  & \\
\Hom_{<\infty}(T(K),\EEx) \arrow{rr}{\text{rec}_T} && H^1(K,\hat{T}_\ell) \\
\end{tikzcd}
\]
We suspect that it is possible to extract Langlands parameters 
from quasicharacter sheaves directly and that the reciprocity map 
can then be be viewed as pairing characters and parameters coming
 from the same quasicharacter sheaf. 
 Results in this paper provide tools for reduction the problem to the case $T=\Gm{K}$. 
 In that case, our suspicions are all but confirmed 
 by the class field theory of Serre-Hazewinkel \citelist{\cite{serre:60a},\cite{serre:61a},\cite{demazure-gabriel:GroupesAlgebriques}*{Appendice}}, especially as revisited in \cite{suzuki-yoshida:12a}. 
 What remains to be done is a careful comparison of 
 quasicharacter sheaves for $\Gm{K}$ and certain local systems on
  the fpqc-site of the pro-quasi-algebraic varieties appearing in \cite{suzuki-yoshida:12a}.

>>>>

Let $\Frob{\Fq}$ denote the geometric Frobenius element in $\Gal(\bFq/\Fq)$ as well as the corresponding automorphism of $\Spec{\bFq}$, and let $\Frob{G} := \id_{G} \times \Frob{\Fq}$ be the Frobenius automorphism of $\bG$.  \David{add definition of Weil groups}

@@

Morphisms in the category of quasicharacter sheaves on $G$, denoted by $\QC(G)$, are defined in the natural way:
\begin{enumerate}
\labitem{(CS.3)}{CS.3} if $\cs{L} = (\gcs{L},\mu,\phi)$ and
  $\cs{L'} = (\gcs{L'},\mu',\phi')$ are quasicharacter sheaves on $G$ then
  a morphism $\cs{L} \to \cs{L}'$ is a map $\alpha : \gcs{L} \to \gcs{L'}$
  of constructible $\ell$-adic sheaves on $\bG$ such that the following diagrams both commute.
  \[
  \begin{tikzcd}[column sep=40]
  \Frob{G}^* \gcs{L} \rar{\Frob{G}^* \alpha} \arrow[swap]{d}{\phi} & \Frob{G}^* \gcs{L'} \dar{\phi'}
  & %\arrow[draw=none]{d}[pos=.4,description]{\text{\normalsize{and}}} 
  & \arrow[swap]{d}{\mu} m^* \gcs{L} \rar{m^* \alpha} & m^* \gcs{L'} \dar{\mu'} \\
  \gcs{L} \rar{\alpha} & \gcs{L'}
  & {} & \LxL \rar{\tight{1}{\alpha\boxtimes \alpha}{1}} & \gcs{L'} \tight{0}{\boxtimes}{0} \gcs{L'}
  \end{tikzcd}
  \]
\end{enumerate}

@@

As mentioned after \cite{deligne:80a}*{Def.~1.1.10}, a Weil sheaf $(\gcs{L},\phi)$ on $G$ may be interpreted as a constructible $\ell$-adic sheaf $\gcs{L}$ on $\bG$ together with an action of the Weil group $\Weil{\Fq}$ on $\gcs{L}$ compatible with the action of $\Gal(\bFq/\Fq)$ on $\bG$. We now review this point of view for later use in Section \ref{sec:basechange}.  If $w = \Frob{\Fq}^n$ we write $w_G$ for $\Frob{G}^n$.  For each such $w$ we define an isomorphism $\varphi(w) : w_G^* \gcs{L}\to \gcs{L}$ by
\[
\varphi(w) \ceq  \phi \circ \Frob{G}^*(\phi) \circ \cdots \circ (\Frob{G}^{n-1})^*(\phi);
\]
these satisfy $\varphi(uv) = \varphi(v) \circ v_G^* \varphi(u)$ for $u,v\in \Weil{\Fq}$, and
we have $\phi = \varphi(\Frob{\Fq})$. 
If we further define $\varphi_1(w) \ceq (w_G)_*(\varphi(w)^{-1})$ then
$\varphi_1(w) : (w_G)_* \gcs{L}\to \gcs{L}$ is an isomorphism and
$\varphi_1(uv) = \varphi_1(u) \circ (u_G)_* \varphi_1(v)$ for $u,v\in \Weil{\Fq}$.
Thus the pair $(\gcs{L},\varphi_1)$ almost satisfies the criteria for an action of $\Weil{\Fq}$ on $\gcs{L}$ compatible with the action of $\Gal(\bFq/\Fq)$ in the sense of \cite{deligne:SGA4.5}*{Expos\'e XIII,~1.1}, failing only because $\Weil{\Fq}$ is profinite \David{Do you mean to say that $\Weil{\Fq}$ is not profinite?}. This fact \David{Which fact?  That we can think about quasicharacter sheaves as an action on $\gcs{L}$?  Or that they fail to fulfill the requirements?} goes to the heart of the notion of a Weil sheaf and is one of the essential aspects of our definition of $\QC(G)$. 
%\end{remark}

@@

%\begin{proposition}\label{tensor}
The category $\QC(G)$ of quasicharacter sheaves on $G$ is rigid symmetric monoidal category \cite{etingof:09a} under the tensor product
$\cs{L} \otimes_{\EE}\cs{L'} \ceq (\gcs{L}\otimes\gcs{L'}, \mu\otimes\mu', \phi\otimes_{\EE}\phi')$
\David{Why only some $\otimes_{\EE}$?}
with duals given by applying the functor 
$\sheafHom(\ - \ ,\EE)$ (sheaf hom) to
conditions \ref{CS.0}, \ref{CS.1} and \ref{CS.2}.
%\end{proposition}
(The category of characters sheaves on $G$ is not abelian, so $\QC(G)$
is not a tensor category in the sense of \cite{deligne:02a}.)
%\begin{proof}\end{proof}
%
This rigid monoidal category structure for $\QC(G)$ puts a group
structure on the set $\QCiso{G}$ of isomorphism classes in
$\QC(G)$. Our main result regarding quasicharacter sheaves on $G$ (Theorem~\ref{thm:snake})
determines $\QCiso{G}$ under the hypothesis that the component group scheme for $G$ is geometrically finitely generated; see Theorem~\ref{thm:snake}. To do this, we use the trace of Frobenius.

@@

\section{Greenberg of N\'eron}
\label{sec:GN}

@@

This paper has shown how quasicharacter sheaves 
on the Greenberg transform of the N\/eron model of $T$ 
determine admissible characters of $T(K)$, and vice versa, functorially. When coupled with the reciprocity map $\text{rec}_T $ for $T$, 
which enjoys the same functorial properties, 
it follows that quasicharacter sheaves for $T$ 
determine Langlands parameters for $T$, and vice versa, functorially.
\[
\begin{tikzcd}
\ & \arrow[swap]{dl}{\trFrob{G}} \QCiso{\GN{T}} \arrow[dashed]{dr}  & \\
\Hom_{<\infty}(T(K),\EEx) \arrow{rr}{\text{rec}_T} && H^1(K,\hat{T}_\ell) \\
\end{tikzcd}
\]
We suspect that it is possible to extract Langlands parameters 
from quasicharacter sheaves directly and that the reciprocity map 
can then be be viewed as pairing characters and parameters coming
 from the same quasicharacter sheaf. 
 Results in this paper provide tools for reduction the problem to the case $T=\Gm{K}$. 
 In that case, our suspicions are all but confirmed 
 by the class field theory of Serre-Hazewinkel \citelist{\cite{serre:60a},\cite{serre:61a},\cite{demazure-gabriel:GroupesAlgebriques}*{Appendice}}, especially as revisited in \cite{suzuki-yoshida:12a}. 
 What remains to be done is a careful comparison of 
 quasicharacter sheaves for $\Gm{K}$ and certain local systems on
  the fpqc-site of the pro-quasi-algebraic varieties appearing in \cite{suzuki-yoshida:12a}.
